### 2024-05-31 index.qmd
May31st, working version

---
title: "ParaChiCSVO2"
titlerunning: "CSVO2"
author:
  - name: Thomas Gavin Carmichael
    orcid: 0009-0008-6849-5333
    corresponding: false
    email: tgcarmichael@outlook.com
    roles:
      - writing - original draft
      - formal analysis
      - methodology
      - validation
      - visualization
    affiliations:
      - ref: 1
      - ref: 2
    degrees:
      - HBSc
  - name: Alexander Rauscher
    orcid: 0000-0002-1961-8252
    email: rauscher@physics.ubc.ca
    corresponding: false
    roles:
      - writing - review & editing
    affiliations:
      - ref: 3
    degrees:
      - PhD
      - MSc
  - name: Ruth E Grunau
    orcid: 0000-0002-5428-9212
    corresponding: false
    email: rgrunau@mail.ubc.ca
    roles:
      - writing - review & editing
      - funding acquisition
    affiliations:
      - ref: 2
      - ref: 3
  - name: Alexander Mark Weber
    orcid: 0000-0001-7295-0775
    corresponding: true
    email: aweber@bcchr.ca
    roles:
      - project administration
      - supervision
      - validation
      - visualization
      - resources
      - methodology
      - formal analysis
      - funding acquisition
      - writing - review & editing
      - conceptualization
      - data curation
      - investigation
    affiliations:
      - ref: 2
      - ref: 3
    degrees:
      - PhD
      - MSc
affiliations:
  - id: 1
    name: The University of British Columbia
    department: Integrated Sciences
    address: 2329 West Mall
    city: Vancouver
    region: BC
    country: Canada
    postal-code: V6T 1Z4
  - id: 3
    name: The University of British Columbia
    department: Pediatrics
    address: 2329 West Mall
    city: Vancouver
    region: BC
    country: Canada
    postal-code: V6T 1Z4
  - id: 2
    department: BC Children's Hospital Research Institute
    name: The University of British Columbia
    address: 938 West 28th Avenue
    city: Vancouver
    state: BC
    country: Canada
    postal-code: V5Z 4H4

keywords:
  - Quantitative Susceptbility Mapping
  - Preterm
  - Newborn
  - Cerebral Venous Oxygen Saturation
abstract: |
  **Background**: Quantitative susceptibility mapping (QSM) is a magnetic resonance imaging (MRI) modality proposed to be a viable method of measuring cerebral oxygenation in neonates given its sensitivity to deoxyhemoglobin, a paramagnetic molecule. During QSM, however, paramagnetic sources can be obscured by opposing diamagnetic sources such as water and myelin. We sought to evaluate whether QSM images alone, or an algorithm that attempts to isolate their paramagnetic components, are more accurate in measuring oxygenation of the major cerebral veins in a cohort of neonates born preterm. Additionally, we aimed to determine whether a difference in oxygenation existed between the major cerebral veins.

  **Methods**: 19 neonates born preterm were scanned on a 3T research MRI at term equivalent age. The protocol included a multi-echo susceptibility-weighted imaging sequence. The acquired imaging data were processed as QSM images to obtain the susceptibility values of the superior sagittal sinus (SSS) and central cerebral veins (CCV). These values were used to calculate the oxygen saturation (SvO2) of the SSS and CCV.  QSM images were subsequently processed to isolate their paramagnetic components. SvO2 values of the SSS and CCV were calculated again from the paramagnetic components.

  **Results**: The mean SvO2 values of the SSS and CCV calculated from QSM images were found to be 72.4% (SD, 3.4%) and 68.7% (SD, 3.5%), respectively. The mean SvO2 values calculated from paramagnetic components were found to be 58.1% (SD, 7.3%) for the SSS and 57.7% (SD, 7.0%) for the CCV.

  **Conclusion**: SSS SvO2 values derived from paramagnetic components agreed well with the existing literature and were closer than the values derived from QSM, however, they displayed greater variability. Although the CCV SvO2 data from QSM aligns more closely with existing literature, it is important to note that the current literature on this topic remains relatively limited in the CCV. Thus, decomposing QSM images into paramagnetic components shows great promise as a method for more accurately measuring cerebral oxygenation in neonates but may require more research to improve precision. Notably, no significant difference in oxygenation was observed between the CCV and the SSS, contrasting with previous studies.

plain-language-summary: |

key-points:
  - 
date: last-modified
bibliography: [ Gavin_Thesis_Ref.bib ]
citation:
  container-title: Unpublished
number-sections: false
notebook-links: false
---

```{r setup, include=FALSE}
# options
knitr::opts_chunk$set(
  # fig.width=8, fig.height=5, 
  # out.width="50%",
  # fig.align="center",
  echo=FALSE, 
  message=FALSE, 
  warning=FALSE
  # cache=TRUE 
  )
set.seed(1234) # reproducible
options(knitr.kable.NA = '') # how kable handles NA
options(reticulate.repl.quiet = TRUE)
```

```{r libraries}
#libraries
library(tidyverse) # ggplot2, dplyr, tidyr, readr, purrr, tibble, stringr, forcats
theme_set( theme_minimal() ) # ggplot theme
library(broom) # for nice summaries
library(knitr) # 
library(kableExtra) # more tables options. Can cause problems
library(Rmpfr)
# library(reticulate) # incorporate Python
# use_virtualenv('./pythonvenv', required = TRUE) # load pythong venv from path
# library(neurobase)
```

```{r}
load("notebooks/results.RData")
```

# Introduction {#sec-intro}

<!-- should be around 5-6 paragraphs. Aim for 460 words -->

# Methods {#sec-data-methods}

The study was approved by the Clinical Research Ethics Board at the University of British Columbia and Children's & Women's Hospital (H21-00655) and written informed consent was obtained from the parent/guardian for each infant.

## Study population

Participant data comes from a previous study [@zhu-etal-cmro2]. Participants consisted of preterm neonates born between 25- and 31-weeks gestational age (GA) who were admitted to the level III NICU at BC Women’s Hospital. Recruitment took place over a span of one year, from February 2021 to January 2022, facilitated by a dedicated research nurse. Parents of eligible infants were approached by the research nurse prior to discharge from the NICU to explain the study objectives and seek their consent for participation. Infants meeting the criteria for inclusion were scanned for the study if they had already been discharged from the NICU, were in stable condition, and had reached a term equivalent age of 37 to 44 weeks GA. However, certain exclusion criteria were applied to ensure the homogeneity and integrity of the study sample. Infants were excluded if there was clinical evidence of a congenital malformation or syndrome, a TORCH infection, or ultrasound evidence of large parenchymal hemorrhagic infarction (\>2 cm, Grade 4 intraventricular hemorrhage).

## Image acquisition

MR imaging was performed on a 3.0 Tesla General Electric Discovery MR750 scanner (scanner software version DV26.0_R03) equipped with a SREE Medical Systems single-channel neonatal head coil (Table \ref{tbl-mri}). The scans were conducted at the BC Children’s MRI Research Facility. Prior to the scanning procedure, subjects were carefully prepared by a research nurse. Swaddling and feeding were used to ensure the comfort and cooperation of the subjects during the scan. Importantly, no sedatives or invasive markers were utilized throughout the procedure. Subjects were placed within a specially designed SREE Medical Systems MRI compatible incubator, which facilitated both safety and motion minimization. Molded foam was strategically positioned around the head and body within the incubator to further restrict subject movement. To protect against potential hearing damage, ear plugs were employed during the scanning process. Additionally, a pulse oximeter was affixed to the subject’s foot to monitor arterial oxygen saturation and heart rate throughout the scan.

```{=tex}
\begin{table}[h]
  \caption{Technical parameters for MR imaging pulse sequences}\label{tbl-mri}%
    \centering
    \begin{tabular*}{\textwidth}{@{}lllll@{}}
    \toprule
        ~ & T1w & T2w & pcASL & SWI \\ \midrule
        Sequence & 3D FSPGR & 3D CUBE & \makecell[tl]{Multi-shot 3D fast \\ spin-echo} & \makecell[tl]{3D spoiled GRE \\ flow-compensated} \\ 
        \makecell[tl]{Phase-encoding \\ direction} & Coronal & Sagittal & Axial & Axial \\ 
        TR (ms) & 7.74 & 2,300 & 4,680 & 30.9 \\ 
        TE (ms) & 2.97 & 66.29 & 10.55 & 5.24 \\ 
        Flip angle & 12$^{\circ}$ & 90$^{\circ}$ & 111$^{\circ}$ & 20$^{\circ}$ \\ 
        FOV (cm) & 20 & 20 & 24 & 25 \\ 
        Acquisition Matrix & 512 x 512 & 256 x 256 & 128 x 128 & 256 x 256 \\ 
        \makecell[tl]{In-plane resolution \\ (mm)} & 0.39 x 0.39 & 0.78 x 0.78 & 1.875 x 1.875 & 0.977 x 0.977 \\ 
        \makecell[tl]{Slice thickness \\ (mm)} & 1 & 1 & 4 & \makecell[tl]{2, reconstructed \\ to 1 with zero \\ filling (ZIP2)} \\ 
        Number of slices & 126 & 106 & 50 & 92 \\ 
        Additional parameters & n/a & n/a & \makecell[tl]{1,450 ms label period; \\ 2,025 ms pulse lable; \\ 24 control-label pairs} & n/a \\ 
        Scan duration & 4 min 39 s & 5 min 1 s & 5 min 26 s & 5 min 29 s \\ \botrule
    \end{tabular*}
    \footnotetext{T1w = T1-weighted; T2w = T2-weighted; pcASL = pseudo-continuous arterial spin labelling; SWI = susceptibility weighted imaging; FSPG = fast spoiled gradient echo; CUBE = General Electric name of sequence, not an acronym; GRE = gradient echo; ZIP2 = through-plane zero filling interpolation}
\end{table}
```
The MRI scan protocol comprised of the following sequences: a T1-weighted scan, a T2-weighted scan, a pseudo-continuous arterial spin labeling (ASL) scan, a multi-echo susceptibility-weighted imaging scan, and a diffusion-weighted imaging (DWI) spin-echo echo planar imaging (EPI) sequence. The DWI sequence was not used for the present study.

## Image analysis

The raw DICOM files acquired from the scanning procedure were converted to NIfTI (Neuroimaging Informatics Technology Initiative) format using Chris Rorden’s \texttt{dcmniix} tool. SWI magnitude data files were then used to create subject-specific brain masks that would not erode the SSS during QSM processing, an issue faced by our group in the past [@liFirstStepNeuroimaging2016]. A step-by-step summary of the pipeline used is shown in @fig-sample.

{{< embed notebooks/Figures.ipynb#fig-graph >}}

First, the fifth echo SWI magnitude file was processed using FSL’s (v. 6.0.7.3) [@woolrichBayesianAnalysisNeuroimaging2009] \texttt{fslroi}, \texttt{fslmaths}, and \texttt{bet} [@smithFastRobustAutomated2002] to create a preliminary brain mask, similar to our previous efforts, which does not contain the SSS. \texttt{Fslroi} was used to isolate the fifth echo of the magnitude data, which was then squared using Fslmaths and the option \texttt{-sqr}. Squaring the magnitude image was found to dramatically improve subsequent brain extraction. The resulting image was then used to create the preliminary brain mask using bet with the options \texttt{-m} and \texttt{-R}. The former flag generated a binary brain mask, while the latter performed a more robust brain centre estimation. The brain mask was then dilated by 7 voxels using \texttt{Fslmaths} and the options \texttt{-kernel boxv} and \texttt{-dilM} in order for the dilated mask to contain the SSS (along with unwanted tissue as well). This mask was then used, along with the phase images, in a MATLAB script for QSM calculation from Christian Kames [@kamesRapidTwostepDipole2018] to produce a preliminary QSM image that contained the SSS, albeit with fairly low signal-to-noise ratio and other unwanted tissue. Given the high contrast in voxel intensity between the SSS and surrounding tissue, the select by intensity tool in \texttt{FSLeyes} [@mccarthyFSLeyes2023] was then used to segment the SSS from the QSM image and create a 3D mask of the selected region. Using \texttt{fslmaths} and the options \texttt{-add} and \texttt{-bin}, the SSS mask was then combined with the original brain mask of the fifth echo. This resulted in a brain mask that contained only brain and SSS signal. Finally, this mask was used in a final QSM post-processing step to create a QSM image that includes the SSS while maintaining a high signal-to-noise ratio, making it suitable to obtain accurate susceptibility values.

STI Suite (v. 3.0) [@liIntegratedLaplacianBased2014], was used to process the final QSM images as it produced the cleanest images without eroding the SSS. The finalized brain mask and the last three echoes of the magnitude and phase images were used in STI Suite along with the following parameters: 0.9766 x 0.9766 x 1 mm$^{3}$ voxel size, 5 ms TE1, 5.3 ms $\Delta$TE, and 77.4 ms sum TE, B0 strength = 3, and B0 direction = [0, 0, 1]. The 3D GRE data option was selected for the phase processing stage, and STAR-QSM was selected for the QSM stage. Finally, the `select by intensity' tool in \texttt{FSLeyes} was then used to semi-automatically make vascular masks of the SSS and CCV from each subject’s QSM image. The vascular masks were used to calculate the mean susceptibility of each subject’s SSS and CCV from their QSM image with \texttt{fslstats}.

To isolate the paramagnetic component of subjects’ QSM data, the χ-separation toolbox28 from the Laboratory for Imaging Science and Technology was used. Each subject’s magnitude and phase SWI data were used along with the following parameters: 0.9766 x 0.9766 x 1 mm$^{3}$ voxel size; TE (s) = [0.005, 0.0102, 0.0155, 0.0207, 0.026]; delta TE (s) = 0.0052; B0 strength = 3; B0 direction = [0, 0, 1]. The mean susceptibility of each subject’s SSS and CCV in their paramagnetic maps was calculated with the same vascular masks used for the QSM images. Sample images showing the magnitude, final QSM, and final paramagnetic component images are shown in @fig-sample.

{{< embed notebooks/Figures.ipynb#fig-sample >}}

Once the mean susceptibility values of the SSS and CCV were obtained from the subjects’ QSM images and paramagnetic maps, venous oxygen saturation (SvO2) was calculated with the following equation [@bergInvestigatingEffectFlow2021]:

$$
SvO_{2} = 1 - \frac{\Delta \chi _{blood} - (\Delta \chi _{oxy} * Hct)}{\Delta \chi _{do} * Hct}
$$ {#eq-svo}

where $\Delta \chi _{blood}$ is the vessel’s measured susceptibility, $\Delta \chi _{oxy}$ is the constant representing the susceptibility changes of oxygenated blood relation to water, $\Delta \chi _{do}$ is the difference in susceptibility between oxygenated and deoxygenated blood, and Hct is the subject’s hematocrit. $\Delta \chi _{oxy}$ was -0.21 * 4$\pi$ ppm as per @portnoyHumanUmbilicalCord2018 and [@sedlacikObtainingBloodOxygenation2007], while $\Delta \chi _{do}$ was -0.03 * 4$\pi$ ppm as per [@weisskoffMRISusceptometryImagebased1992]. Subjects’ Hct for the day of the scan was calculated using a four-parameter Weibull function with previously measured values while still in the NICU.

## Statistical analysis

Statistical analysis of the acquired data was performed using R and RStudio (v. 2023.09.1 Build 494) [@rcoreteamLanguageEnvironmentStatistical2022; @rstudioteamRStudioIntegratedDevelopment]. A Wilcoxon signed-rank test was used to determine if a significant difference (p <0.05) existed between the vessel-specific SvO2 values determined through QSM and those determined from the paramagnetic maps processed with $\chi$-separation. This was also performed for the raw $\chi$ values obtained through both methods.

# Results {#sec-results}

A total sample size of 19 infants were scanned, with a mean gestational age of 28.8 weeks (SD, 1.7) and a mean birth weight of 1276.1 grams (SD, 294.9). A comprehensive summary of neonatal characteristics, including additional demographic and clinical data, is provided in Table \ref{tbl-chistats} for reference.

```{=latex}
\begin{table*}[h]
  \caption{Technical parameters for MR imaging pulse sequences}\label{tbl-chistats}%
    \centering
    \begin{tabular*}{\textwidth}{@{}llllll@{}}
    \toprule
        Region & Measure & QSM & \makecell[tl]{Paramagnetic \\ map} & p-value & 95\% CI \\ \midrule
        \multirow{ 2}{*}{SSS} & Chi (ppm) & `r round(mean(newdata$Gavin_SSS_Chi),2)` \pm `r round(sd(newdata$Gavin_SSS_Chi),2)` & `r round(mean(newdata$Chisep_SSS_Chi),2)` \pm `r round(sd(newdata$Chisep_SSS_Chi),2)` & `r signif(SSSChi.t$p.value,3)` & `r round(SSSChi.t$conf.low,2)`, `r round(SSSChi.t$conf.high,2)`\\
        & SvO$_{2}$(\%) & `r round(mean(newdata$Gavin_SSSVein_CSvO2*100),2)` \pm `r round(sd(newdata$Gavin_SSSVein_CSvO2*100),2)` & `r round(mean(newdata$Chisep_SSSVein_CSvO2*100),2)` \pm `r round(sd(newdata$Chisep_SSSVein_CSvO2*100),2)` & `r signif(SSSSvO2.t$p.value,3)` & `r round(SSSSvO2.t$conf.low*100,2)`, `r round(SSSSvO2.t$conf.high*100,2)`\\
        \multirow{ 2}{*}{CCV} & Chi (ppm) & `r round(mean(newdata$Gavin_IntVein_Chi),2)` \pm `r round(sd(newdata$Gavin_IntVein_Chi),2)` & `r round(mean(newdata$Chisep_IntVein_Chi),2)` \pm `r round(sd(newdata$Chisep_IntVein_Chi),2)` & `r signif(CCVChi.t$p.value,3)` & `r round(CCVChi.t$conf.low,2)`, `r round(CCVChi.t$conf.high,2)`\\
        & SvO$_{2}$(\%) & `r round(mean(newdata$Gavin_IntVein_CSvO2*100),2)` \pm `r round(sd(newdata$Gavin_IntVein_CSvO2*100),2)` & `r round(mean(newdata$Chisep_IntVein_CSvO2*100),2)` \pm `r round(sd(newdata$Chisep_IntVein_CSvO2*100),2)` & `r signif(CCVSvO2.t$p.value,3)` & `r round(CCVSvO2.t$conf.low*100,2)`, `r round(CCVSvO2.t$conf.high*100,2)`\\
        \botrule
    \end{tabular*}
    \footnotetext{}
\end{table*}
```

<!-- ::: {#tbl-1} -->

<!-- ```{=latex} -->
<!-- % Or whatever actual you want, includegraphics, whatever.  -->
<!-- % (_not_ \begin{table}, just the content of it) -->
<!-- \begin{tabular}{lll} -->
<!-- \end{tabular} -->
<!-- ``` -->

<!-- This is a table caption. -->

<!-- ::: -->

<!-- See @tbl-1 -->

See @fig-regionplot

{{< embed notebooks/gavin_thesis_markdown.qmd#fig-regionplot >}}

# Conclusion {#sec-conclusion}

# References {.unnumbered}

::: {#refs}
:::