### 2024-05-31 notebooks/gavin_thesis_markdown.qmd

---
title: "gavin_thesis_R"
author: "Gavin"
date: "2024-04-30"
output: html_document
---

```{r}
#| label = "setup",
#| include = FALSE
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(ggstatsplot)
library(rstatix)
library(patchwork)
```

Load main dataframe (change directory as needed)
*note: columns titled with with "Gavin_x_x" contain QSM data
```{r}
newdata <- read.csv("newdata.csv")
```

Dataframes for figures
```{r}
#| echo = FALSE

#QSM chi data - SSS vs. CCV
comp_legit_starfsl <- data.frame(
  Chi = c(newdata$Gavin_SSS_Chi, newdata$Gavin_IntVein_Chi),
  Vein = c(rep("Superior Sagittal Sinus", 19), rep("Central Cerebral Veins", 19))
)

#QSM CSvO2 data - SSS vs. CCV
comp_SSS_IntVein_CSvO2 <- data.frame(
  CSvO2 = c(newdata$Gavin_IntVein_CSvO2, newdata$Gavin_SSSVein_CSvO2),
  Vein = c(rep("Central Cerebral Veins", 19), rep("Superior Sagittal Sinus", 19))
)

#Chisep chi data - SSS vs. CCV
comp_chisep_chi <- data.frame(
  Chi = c(newdata$Chisep_IntVein_Chi, newdata$Chisep_SSS_Chi),
  Vein = c(rep("Central Cerebral Veins", 19), rep("Superior Sagittal Sinus", 19))
)

#Chisep CSvO2 data - SSS vs. CCV
comp_chisep_CSvO2 <- data.frame(
  CSvO2 = c(newdata$Chisep_IntVein_CSvO2, newdata$Chisep_SSSVein_CSvO2),
  Vein = c(rep("Central Cerebral Veins", 19), rep("Superior Sagittal Sinus", 19))
)

#QSM SSS chi vs. chisep SSS chi
comp_SSS_chi <-  data.frame(
  "Measured Susceptibility" = 
    c(newdata$Gavin_SSS_Chi, newdata$Chisep_SSS_Chi),
  Method = c(rep("QSM", 19), rep("Paramagnetic Component", 19))
)

#QSM CCV chi vs. chisep CCV chi
comp_CCV_chi <- data.frame(
  Chi = 
    c(newdata$Gavin_IntVein_Chi, newdata$Chisep_IntVein_Chi),
  Method = c(rep("QSM", 19), rep("Paramagnetic Component", 19))
)

#QSM SSS SvO2 vs. chisep SSS SvO2
comp_SSS_SvO2 <- data.frame(
  "Oxygen Saturation" = 
    c(newdata$Gavin_SSSVein_CSvO2, newdata$Chisep_SSSVein_CSvO2),
  Method = c(rep("QSM", 19), rep("Paramagnetic Component", 19))
)

#QSM CCV SvO2 vs. chisep CCV SvO2
comp_CCV_SvO2 <- data.frame(
  "Oxygen Saturation" = 
    c(newdata$Gavin_IntVein_CSvO2, newdata$Chisep_IntVein_CSvO2),
  Method = c(rep("QSM", 19), rep("Paramagnetic Component", 19))
)
```

Wilcox test for p-values and 95% CI - table 3.2
```{r}
#SSS Chi
wilcox.test(newdata$Gavin_SSS_Chi, newdata$Chisep_SSS_Chi,
            paired = TRUE, conf.int = TRUE)

#SSS SvO2
wilcox.test(newdata$Gavin_SSSVein_CSvO2, newdata$Chisep_SSSVein_CSvO2,
            paired = TRUE, conf.int = TRUE)

#CCV Chi
wilcox.test(newdata$Gavin_IntVein_Chi, newdata$Chisep_IntVein_Chi, paired = TRUE,
            conf.int = TRUE)

#CCV SvO2
wilcox.test(newdata$Gavin_IntVein_CSvO2, newdata$Chisep_IntVein_CSvO2,
            paired = TRUE, conf.int = TRUE)
```

```{r}
#SSS Chi
SSSChi.t <- tidy(t.test(newdata$Gavin_SSS_Chi, newdata$Chisep_SSS_Chi,
            paired = TRUE, conf.int = TRUE))

#SSS SvO2
SSSSvO2.t <- tidy(t.test(newdata$Gavin_SSSVein_CSvO2, newdata$Chisep_SSSVein_CSvO2,
            paired = TRUE, conf.int = TRUE))

#CCV Chi
CCVChi.t <- tidy(t.test(newdata$Gavin_IntVein_Chi, newdata$Chisep_IntVein_Chi, paired = TRUE,
            conf.int = TRUE))

#CCV SvO2
CCVSvO2.t <- tidy(t.test(newdata$Gavin_IntVein_CSvO2, newdata$Chisep_IntVein_CSvO2,
            paired = TRUE, conf.int = TRUE))
```

Figure 3.1
*note: "****" in panels were added outside of R
```{r}
#| echo = FALSE
m1 <- ggwithinstats(
  data = comp_CCV_chi,
  x = Method,
  y = Chi,
  results.subtitle = FALSE,
  title = "(A)",
  subtitle = "Central Cerebral Veins",
  type = "nonparametric",
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid")
)

m2 <- ggwithinstats( 
  data = comp_SSS_chi,
  x = Method,
  y = Measured.Susceptibility, 
  ylab = "Chi",
  results.subtitle = FALSE,
  title = "",
  subtitle = "Superior Sagittal Sinus",
  type = "nonparametric", 
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid")
)

m3 <- ggwithinstats(
  data = comp_CCV_SvO2,
  x = Method,
  y = Oxygen.Saturation,
  ylab = "SvO2",
  results.subtitle = FALSE,
  title = "(B)",
  subtitle = "Central Cerebral Veins",
  type = "nonparametric",
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid")
)

m4 <- ggwithinstats(
  data = comp_SSS_SvO2,
  x = Method,
  y = Oxygen.Saturation,
  pairwise.display = "all",
  ylab = "SvO2",
  results.subtitle = FALSE,
  title = "",
  subtitle = "Superior Sagittal Sinus",
  type = "nonparametric",
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid")
)

combine_plots(
  plotlist = list(m1, m2, m3, m4),
  plotgrid.args = list(nrow = 2)
)
```

Figure 3.2
*note: "****" in panels (a) and (b) were added outside of R
```{r}
#| echo = FALSE
p1 <- ggwithinstats(
  data = comp_legit_starfsl,
  x = Vein,
  y = Chi,
  pairwise.display = "all",
  results.subtitle = FALSE,
  xlab = "Region",
  ylab = "Chi",
  title = "(A)",
  type = "nonparametric",
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid"),
  point.args = list(color = "#D95F02") 
)

p2 <- ggwithinstats(
  data = comp_SSS_IntVein_CSvO2,
  x = Vein,
  y = CSvO2,
  pairwise.display = "all",
  results.subtitle = FALSE,
  xlab = "Region",
  ylab = "SvO2",
  title = "(B)",
  type = "nonparametric",
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid"),
  point.args = list(color = "#D95F02")
)

p3 <- ggwithinstats(                     #boxplot of chi comparison between regions
  data = comp_chisep_chi,
  x = Vein,
  y = Chi, 
  pairwise.display = "all",
  results.subtitle = FALSE,
  xlab = "Region",
  ylab = "Chi",
  title = "(C)",
  subtitle = "                                                n.s.",
  type = "nonparametric", 
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid"),
  point.args = list(color = "#1B9E77")
)

p4 <- ggwithinstats(                    #boxplot of svo2 comparison between regions
  data = comp_chisep_CSvO2,
  x = Vein,
  y = CSvO2, 
  pairwise.display = "all",
  results.subtitle = FALSE,
  xlab = "Region",
  ylab = "SvO2",
  title = "(D)",
  subtitle = "                                                n.s.",
  type = "nonparametric",
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid"),
  point.args = list(color = "#1B9E77")
)

combine_plots(
  plotlist = list(p1, p2, p3, p4),
  plotgrid.args = list(nrow = 2)
)
```

```{r}
save.image('results.RData')
```

```{r}
 p1 <- ggwithinstats(
  data = comp_legit_starfsl,
  x = Vein,
  y = Chi,
  results.subtitle = FALSE,
  type = "p",
  pairwise.display = "all",
  bf.message = FALSE,
  xlab = "Region",
  ylab = "Chi",
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid"),
  point.args = list(color = "#D95F02")) + 
  ggplot2::annotate("text", x=1.5, y=(max(comp_legit_starfsl$Chi)*1.05), size=4, label = "***")

p2 <- ggwithinstats(
  data = comp_SSS_IntVein_CSvO2,
  x = Vein,
  y = CSvO2,
  results.subtitle = FALSE,
  type = "p",
  pairwise.display = "all",
  bf.message = FALSE,
  xlab = "Region",
  ylab = "SvO2",
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid"),
  point.args = list(color = "#D95F02")
) + 
  ggplot2::annotate("text", x=1.5, y=(max(comp_SSS_IntVein_CSvO2$CSvO2)*1.05), size=4, label = "***")

p3 <- ggwithinstats(                     #boxplot of chi comparison between regions
  data = comp_chisep_chi,
  x = Vein,
  y = Chi, 
  results.subtitle = FALSE,
  pairwise.display = "all",
  bf.message = FALSE,
  xlab = "Region",
  ylab = "Chi",
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid"),
  point.args = list(color = "#1B9E77")
) +
  ggplot2::annotate("text", x=1.5, y=(max(comp_chisep_chi$Chi))*1.05, size=4, label = "n.s.")

p4 <- ggwithinstats(                    #boxplot of svo2 comparison between regions
  data = comp_chisep_CSvO2,
  x = Vein,
  y = CSvO2, 
  results.subtitle = FALSE,
  pairwise.display = "all",
  bf.message = FALSE,
  xlab = "Region",
  ylab = "SvO2",
  type = "nonparametric",
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid"),
  point.args = list(color = "#1B9E77")
) +
  ggplot2::annotate("text", x=1.5, y=(max(comp_chisep_CSvO2$CSvO2)*1.05), size=4, label = "n.s.")

(patch1 <- (p1 + p2)/(p3 + p4))
# patch1
# ggsave('figure.png', patch1)
```

```{r}
#| label: fig-regionplot
#| fig-width: 8
#| fig-height: 5
#| fig-cap: "rmarkdown test figure"
patch1
```